\section{Related Work}
\label{sec:related}

We already discussed prior work on OLED power modeling in
\S\ref{sec:non_linear}. Below, we discuss previous work on optimizing
OLED power draw via changing screen colors.

{\bf Color remapping.}
Several studies explore the simple idea of selectively dimming portions
of the OLED display panel according to the user's focus of attention
(FOA)~\cite{tan:ubicom13,chen:hotpower14} which are generally hard to
predict accurately.
%
More recent work focus on color remappings that exploit
approximation of original color content that have lower OLED display
power draw. In particular, application-specific color remapping
modifies individual applications such as games~\cite{anand:mobisys11},
web browsers~\cite{dong:ispled09,li:esec15} and web
servers~\cite{li:icse2014} to carry our pre-coded color remappings,
while system-level color remapping~\cite{crayon:eurosys16}
transparently generates color mappings (with controllable
quality degradation) for dynamically generated images before
displaying them.
% on the screen.

{\bf Dark mode.}
%  Many have studied the color transformations
%  ~\cite{darkmode:stanley2016crayon} to reduce number of bright pixels
%  on the display by using the content on the screen to create dark modes
%  by applying color transformation ~\cite{banman2019smartnight}
%  ~\cite{dong:2011:chameleon} ~\cite{darkmode:dong2009power} These
%  techniques are based on heuristic and may be computationally
%  expensive.
Dark mode can be viewed as a special case
of color transformation that reduces screen power
by explicitly flipping white-themed background colors to dark-themed colors.
Both Google with Android Q~\cite{darkmode:androidQ} and
Apple with iOS 13~\cite{darkmode:iOS13_1,darkmode:iOS13_2} have added support
for dark mode.  Google is also working on Force Dark
Mode~\cite{darkmode:forcedark} which will apply dark mode on apps
that do not support the mode yet.
%
Dark mode is also being adopted in non-mobile platforms such as
laptops and in web browsers~\cite{darkmode:web_browsers} which
perform color inversion as most web pages have light
background with black text.
%  This is turns out to be tricky as in web browser especially graphics once has
%  to be handled such that they are not affected.
Chrome ~\cite{darkmode:chrome} has added flag settings to force dark mode
and intelligent implementation to find and not invert image content.
